---
title: "Practical Machine Learning Course Project"
author: "Edward Gañan"
date: "Sunday, May 24, 2015"
output: html_document
---

This project is about human activity recognition and is concerned on how well an activity was performed by the wearer. the dataset was obteined from  http://groupware.les.inf.puc-rio.br/har where 6 participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways while using accelerometers on the belt, forearm, arm, and dumbell. The output was clasified in classes A, B, C, D y E, where Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes. The goal of this project is to try to predict output class based on accelerometers measures

##Getting an cleanig data

The data are downloaded using download.file function, we have 19622 records and   160 variables
```{r,echo=TRUE}
pml_training<-read.csv("pml-training.csv",na.strings=c("","#DIV/0!","NA"),
                       stringsAsFactors=F)
```

takes variables only related to belt, forearm, arm, dumbell
```{r,echo=TRUE}
pml_training<-pml_training[,-c(1:7)]
```


remove variables with a lots NA. We ended with 52 predictors
```{r,echo=TRUE}
na.fn<-function(x){
  y<-prop.table(table(is.na(x)))
  r<-unname(y[names(y)==T])
  if(length(r)==0) r<-0
  r
}

var.nas<-sapply(pml_training,na.fn)
var.nas<-names(var.nas[var.nas>0.1])
pml_training<-subset(pml_training,select=names(pml_training) %in% var.nas ==F)
```

finally, we validate that we have complete cases
```{r,echo=TRUE}
table(complete.cases(pml_training))
```


##Somme exploratory anlysis
We have approximately balanced dataset
```{r,echo=TRUE,fig.height=3.2,fig.width=6}
dat<-as.data.frame(table(pml_training$classe))
names(dat)[1]<-"classe"
library(ggplot2)
p<-ggplot(dat,aes(x=classe,y=Freq))+geom_bar(stat="identity")+theme_bw()
p
```


We see that all variables have an important variability
```{r,echo=TRUE}
library(caret)
nsv<-nearZeroVar(pml_training,saveMetrics=T)
nsv
#All are FALSE
```

Reduce dimensionality to analysis classes. 12 components are required to represent 80% of variability
```{r,echo=TRUE}
pca<-princomp(pml_training[,-53],cor=T)
pca.scores<-as.data.frame(pca$scores)
pca.scores$classe<-pml_training$classe
p<-ggplot(data=pca.scores,aes(x=Comp.1,y=Comp.2,color=classe))+
    geom_point()
p<-p+theme_bw()
p
```


##Adjust prediction model
For predcition we use random forest because can consider interacction between variables and its good performance in accuracy
```{r,echo=TRUE}
pml_training$classe<-as.factor(pml_training$classe)
library(randomForest)
set.seed(12568)
modelRF<-randomForest(classe~.,ntree=1000,data=pml_training)
modelRF
```

OOB it's a good estimate of out of sample error and we are getting a low error prediction. Also, random forest already consider Bagging for prediction outcome.




